---
title: "P1"
author: "Simran Tejay"
date: '2019-03-25'
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
DataRead:
```{r}
data <- read.csv("Cost_of_living_index.csv", header=TRUE)
```

## Multiple Linear Regression

```{r}
fit <- lm(data$Rank~data$Cost.of.Living.Index+data$Rent.Index+data$Cost.of.Living.Plus.Rent.Index+data$Groceries.Index+data$Restaurant.Price.Index+data$Local.Purchasing.Power.Index)
summary(fit)
```
By performing a Multiple Linear Regression on this dataset, with "Rank" being the response variable, we are able to determine how much of an influence each predictor has on the response variable "Rank". This measure of significance is expressed in the above summary output of the data where we can see that the "Rank" increases relative to the estimate values for the predictor variables. Using this method of analysis we can determine that, for instance, as the "Cost of Living Index" increases, the rate at which rank will increase by is 53.49352. The influence of the "Rent Index" is also very similar with a estimate of 55.09453. An interesting predictor that has a negative correlation with respect to rank is the "Cost of Living Plus Rent Index" where the rank of the city increases as the predictor decreases at a rate estimate of -117.73226. Lastly the "Groceries Index", "Restaurant Price Index", and "Local Purchasing Power Index" predictors are less significant predictors with respect to the ranking of countries compared to the "Cost of Living Index" and "Rent Index". Therefore, the cost of living index, rank index and cost of living plus rent index have the greatest influence on the rank of the countries within this dataset and are the most significant predictors with respect to the rank of a country. While the remaining predictors are far less significant in comparison. We are able to draw this inference since our Adjusted R-squared is equal to 0.9345 which tells us there is minor variation as the independent variables explain our response variable. 


## Regression Tree
```{r}
library(tree)
library(gclus)
data1 <- data.frame(data)
rank <- data1$Rank
city <- data1$City
cost_of_living <- data1$Cost.of.Living.Index
rent <- data1$Rent.Index
living_plus_rent <- data1$Cost.of.Living.Plus.Rent.Index 
groceries <- data1$Groceries.Index
restaurant <- data1$Restaurant.Price.Index
purchasing_power <- data1$Local.Purchasing.Power.Index

simtree <- tree(rank~cost_of_living+rent+living_plus_rent+groceries+restaurant+purchasing_power)
summary(simtree)
plot(simtree)
text(simtree, pretty=0)
```
```{r}
#set.seed(450)
#cvsimtree <- cv.tree(as.matrix(simtree), FUN=prune.tree, K=10)
#plot(cvsimtree, type="b")
#newtree <- prune.tree(as.matrix(simtree), best=7)
#plot(newtree)
#text(newtree, pretty=0)
```



